# Chapter 6 Selected Exercises

```{r}
library(tidyverse)
library(rstanarm)
```

## Exercise 6.2
Simulate n data points from y = a + bx + error,  with x uniformly sampled from (0,100).  error has standard deviation sigma (parameter).

```{r}

simulate_data <- \(a, b , n, sigma) {
  meas_err <- rnorm(n,0,sigma)
  tibble(x = runif(n,0,100),
  y = a + b*x + meas_err)
  
}

sim_fit_plot <- \(a, b , n, sigma, quiet = FALSE){
 data <- simulate_data(a,b,n,sigma) 
 fit <- stan_glm(y ~  x,data = data, refresh=0)
 a_hat = coef(fit)[1]
 b_hat = coef(fit)[2]
 if(!quiet){
   print(fit)
   (ggplot(data, mapping = aes(x = x, y = y)) +
   geom_point() + 
   geom_abline(slope= b_hat, intercept = a_hat)) |> print()
 }
 list(data= data, fit = fit)
}

test <- sim_fit_plot(10,.5,100,30)

```

## Exercise 6.3

Lets pick a = 0, b= 1, and sigma = 30 as the case so there is plenty of noise.

```{r}
test1 <- sim_fit_plot(0,1,100,30)
```
```{r}
test2 <- sim_fit_plot(0,1,10,30)
```


```{r}
test2 <- sim_fit_plot(0,1,3,30)
```

We can see that with less points, the errors on the estimates are larger as are the estimated uncertainties. Note that with n=2 the fit has trouble (since we are estimating two parameters)

## Exercise 6.4

This approach is not very pretty
```{r}
ns <- seq(10,100)
a <- 0
b <- 1
sigma <- 30
results <- tibble(n =ns, a_hat= rep(0, length(ns)), b_hat = a_hat, a_mad_sd = a_hat, b_mad_sd = b_hat)
for(i in seq_along(ns)){
  fit <- sim_fit_plot(a,b,ns[[i]],sigma, quiet=TRUE)$fit
  results$a_hat[[i]] <- coef(fit)[1]
  results$b_hat[[i]] <- coef(fit)[2]
  results$a_mad_sd[[i]] <-fit$ses[1]
  results$b_mad_sd[[i]] <- fit$ses[2]
  
}
```

Now to plot it

```{r}
results |> pivot_longer(cols = c('a_hat','b_hat','a_mad_sd','b_mad_sd'), names_to = "param") |>
ggplot(aes(x=n,y=value)) + facet_wrap(~param, scales="free_y" ) + geom_line()

```

  As n increases , the errors get smaller. I could probably plot this better. Also look at vs sqrt(n)
  
```{r}

results |> pivot_longer(cols = c('a_hat','b_hat','a_mad_sd','b_mad_sd'), names_to = "param") |>
ggplot(aes(x=sqrt(n),y=value)) + facet_wrap(~param, scales="free_y" ) + geom_line()

```
Not quite sqrt(n) as I would expect? 

## Exercise 6.6

## Exercise 6.8