# Selected Exercises
```{r}
library(tidyverse)
library(rstanarm)
library(ggplot2)
library(rstan)
library(brms)
```

## 12.6

### a) Scatter plot of Mort vs NOX

```{r}
pollution <- read_csv("Pollution/data/pollution.csv",
                      show_col_types =  FALSE)

ggplot(data = pollution, aes(x=nox, y = mort)) + geom_point()
```
I doesn't appear that a linear model will work very well. To my eye it looks exponential for small nox and then there are a bunch of outliers.  Lets try a fit:

```{r}
fit <- stan_glm(mort ~ nox, data = pollution, refresh =0)
fit
```

```{r}
ggplot(data = pollution, aes(x=nox, y = mort)) + geom_point() +
  geom_abline(intercept= fit$coefficients[1], slope = fit$coefficients[2])
```

As expected the outliers drive the fit. The coefficient for nox is consistent with zero.


Plotting the residuals:
```{r}
pollution$residual <- pollution$mort - predict(fit)
ggplot(data = pollution) + geom_point(aes(x=nox, y= residual))
```

Explains none of the variance.

###  b) Try transformations

I suspect that the log of the nox will look more linear, and to make it easier to interpret lets also do the log of the mort. 

```{r}
ggplot(data = pollution, aes(x=log(nox), y = log(mort))) + geom_point()

```


On this scale there is a clear linear relationship except I do worry about the points at the right.


```{r}
fit2 <- stan_glm(log(mort) ~ log(nox), data = pollution, refresh =0)
print(fit2, digits =4)
```

Now at least wee see a 'significant' positive effect.


```{r}
ggplot(data = pollution, aes(x=log(nox), y = log(mort))) + geom_point() +
  geom_abline(intercept= fit2$coefficients[1], slope = fit2$coefficients[2])
```

Or on the original scale:

```{r}

fit_line = tibble(nox = 0:300 )
fit_line$mort = exp(predict(fit2, newdata = fit_line))
ggplot(data = pollution, aes(x=nox, y = mort)) + geom_point() +
  geom_line(data = fit_line, mapping = aes(x=nox,y=mort))
```


Plotting the residuals:
```{r}
pollution$residual <- log(pollution$mort) - predict(fit2)
ggplot(data = pollution) + geom_point(aes(x=nox, y= residual))
```

Compare the loo's, using the correction to the Jacobian from the text which I do not understand fully ;)

```{r}
loo1 = loo(fit)
loo2 = loo(fit2)
loo2_with_jacobian <- loo2
# just copied from the book.... 
loo2_with_jacobian$pointwise[,1] <- loo2_with_jacobian$pointwise[,1] - log(pollution$mort)
loo_compare(loo1,loo2_with_jacobian)
```

Fit two is better, just barely significantly though.

```{r}
c(median(bayes_R2(fit)),median(bayes_R2(fit2)))
```

Fit2 does explain more of the variance at least!

### Outliers (not part of the excercise)

That outlier is the pulling the fit, a later chapter will show how to do robust fit using brms, so I read ahead to try it out:

```{r}
fit3 <- brm(log(mort) ~ log(nox), data = pollution, family = 'student', refresh=0)
print(fit3, digits = 4)
```


```{r}
fit3_fit <- fixef(fit3)
```


```{r}
ggplot(data = pollution, aes(x=log(nox), y = log(mort))) + geom_point() +
  geom_abline(intercept= fit3_fit[1], slope = fit3_fit[2])
```


This move the slope up a bit and de-emphasized the outliers as desired.

 
 
```{r}
loo3 = loo(fit3)
loo3_with_jacobian <- loo3
# just copied from the book.... 
loo3_with_jacobian$pointwise[,1] <- loo3_with_jacobian$pointwise[,1] - log(pollution$mort)
loo_compare(loo1,loo2_with_jacobian, loo3_with_jacobian)
```

fit3 is significantly better then the original fit, but not from fit2.

Note we can compare directly loo2 and loo3:

```{r}
loo_compare(loo2,loo3)
```

So the 'robust'fit is only marginally better according to loo, which makes sense do the dominance of the noise.

```{r}
median(bayes_R2(fit2))
```
   
### c) How do we interpret the coefficent on the slope?

The model is: `log(mort) = 0.016*log(nox) + intercept`

or `mort = exp(intercept)*nox^(0.016)`

The slope here give the exponent in the power law., the coefficient is the exponent `c` in:

`mort ~ nox^c'  

for example, each doubling of nox produces about a 1.3 % increase in mortality. 
