## Process data for presentation. 
process nutrition data as an example

```{r}
library(dplyr)
library(tidyr)
library(rstanarm)
nutrition <- readr::read_csv('obesity.csv')
exercise <- nutrition |>
           filter(QuestionID == 'Q036' | QuestionID == 'Q046')  |>
           filter(YearStart == 2015 & StratificationCategory1 == 'Total')  |>
           select(c(LocationAbbr,Question, QuestionID, Data_Value)) |>
           mutate(Variable = if_else(QuestionID == 'Q036', 'obesity','strengthTraining')) |>
           pivot_wider(id_cols = LocationAbbr, names_from = Variable, values_from = Data_Value) |> 
           filter(!(LocationAbbr %in% c('US','GU','PR'))) |>
           rename(state = LocationAbbr)

```

```{r}
library(ggplot2)

ggplot(exercise, aes(x= strengthTraining, y = obesity)) +  geom_text(aes(label= state))

```



### Export data for use in bookdown

```{r}
readr::write_csv(exercise, '2016Exercise.csv')
```


## Selected Exercises  

### 7.3 Fake Data and wrong model

```{r}
n <- 100
a <- 0
b <- -.02
c <- .05
x <- runif(n,0,50)
y <- a + b*x + c* x^2 + rnorm(n,0,3)
fake <- tibble(x,y)
p <- ggplot(fake) + geom_point(aes(x=x,y=y))
p
```
```{r}
fit = stan_glm(y ~x, data = fake, refresh= 0)
print(fit, detail=FALSE)
```
```{r}
p + geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2])
```

Note that the fit is 'best' in that it best approximates the data, the best we can do with a straight line.   even though the data is not a line and the residuals still show a pattern.  

```{r}
fake |> mutate(residual = coef(fit)[1] + x*coef(fit)[2] - y) |>
     ggplot() + geom_point(aes(x=x, y = residual))
```

```{r}
fit = stan_glm(y ~ x + I(x^2), data = fake, refresh= 0)
print(fit, detail=FALSE,digits=3)
```

```{r}
fake |> mutate(poly_fit = coef(fit)[1] + coef(fit)[2]*x + coef(fit)[3]*x^2) |> ggplot() + 
geom_point(aes(x=x, y=y)) +
geom_line(aes(x=x, y= poly_fit))
```

Using a nonlinear fit captures the underlying  generating process well.


### 7.4 (and sort of 7.1)

Using the health data above:
```{r}
fit = stan_glm(obesity ~ strengthTraining, data = exercise, refresh= 0)
print(fit, detail=FALSE)
ggplot(exercise) + geom_point( aes(x= strengthTraining, y = obesity))  +
                   geom_abline(aes(slope = coef(fit)[2], intercept = coef(fit)[1]))  + 
                   geom_text( aes(x= strengthTraining, y = obesity, label=state), hjust=1.2, vjust=1, alpha=0.6) +
                   geom_text( x = 24, y = 25, label = glue::glue('obesity = {round(coef(fit)[1],1)} - {-round(coef(fit)[2],1)} strengthTraining '))
```

So for a 'new' state (or same state but perhaps after a program of promoting strength training) with 35% participation we would predict:

```{r}
61.9 - 1.1*35
```

23% obesity +- 2.5%.  

What is the probability that the obesity will be more then 25%?

```{r}
1-pnorm(25,23,2.5)
```

(Using normal approximation)