# Chapter 14, selected exercises

```{r}
library(rstanarm)
library(tidyverse)

```

## Exercise 14.3 

Graphing Logistic Regressions

Graphing logistic regressions: The well-switching data described in Section 13.7 are in the
folder Arsenic.

### Fit (a)

(a) Fit a logistic regression for the probability of switching using log (distance to nearest safe
well) as a predictor.

```{r, message=FALSE}
wells <- read_csv('wells.csv')
wells <- wells |> mutate(log_dist = log(dist))
summary(wells)
```
```{r}
fit_a <- stan_glm(switch ~ log(dist), data = wells,family = binomial(link = "logit"), refresh = 0)
print(fit_a, digits=3)
```

*I wonder if I should center log distance?* 

### (b) Make a graph similar to Figure 13.8b 

Display Pr(switch) as a function of distance to nearest safe well, along with the data.

Borrowed from tidyros example 

```{r}
v <- 
  tibble(
    dist = seq(from=min(wells$dist), to = max(wells$dist), length.out=100),
    .pred = predict(fit_a, type = "response", newdata = tibble(dist))
  )

wells %>% 
  ggplot(aes(dist)) +
  stat_ydensity(
    aes(y = switch, group = switch),
    width = 0.25,
    draw_quantiles = c(0.25, 0.5, 0.75),
    scale = "count"
  ) +
  geom_line(aes(y = .pred), data = v) +
  coord_cartesian(ylim = c(-0.125, 1.125)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(50)) +
  labs(
    title = "Probability of household switching to new well by distance",
    subtitle =
      "Voilin plots represent density of those do did and did not switch",
    x = "Distance to the closest known safe well (meters)",
    y = "Probability of household switching"
  )
```

### Look at it on log scale, vs binned data

```{r}
binned <- wells |>  mutate(bin = cut_interval(log_dist,n = 10)) |>
        group_by(bin) |> 
             summarize(
             dist = median(dist),
             p_switch = mean(switch),
             p_se = sd(switch) / sqrt(n()) 
             )
             
 

binned |> 
  ggplot() +
  geom_errorbar(aes(x= log(dist), ymin = p_switch - p_se,ymax = p_switch + p_se)) +
  geom_line(aes(x = log(dist), y = predict(fit_a, type = "response")),data=wells, color='red') +
  geom_point(aes(x=log(dist), y= p_switch))
```



### (c) Make a residual plot and binned residual plot as in Figure 14.8.
I am only going to show binned residual plot

This code borrowed from the tidy-ros examples:

```{r}
plot_binned_residuals <- function(data) {
  data %>% 
    ggplot(aes(x_mean)) +
    geom_hline(yintercept = 0, color = "white", size = 2) +
    geom_line(aes(y = 2 * y_se), color = "grey60") +
    geom_line(aes(y = -2 * y_se), color = "grey60") +
    geom_point(aes(y = y_mean)) +
    labs(y = "Average residual")
}

binned_means <- function(x, y, n_bins = 40) {
  tibble(x, y) %>% 
    mutate(bin = cut_number(x, n = n_bins)) %>% 
    group_by(bin) %>% 
    summarize(
      x_mean = mean(x),
      y_mean = mean(y),
      y_se = sd(y) / sqrt(n())
    )
}
```

```{r}
v <- 
  binned_means(
    x = wells$dist,
    y = residuals(fit_a)
  )

v %>% 
  plot_binned_residuals() +
  labs(
    title = "Binned residuals by distance to nearest safe well",
    x = "Distance to nearest safe well"
  )
```


### (d) Compute the error rate of the fitted model and compare to the error rate of the null model.

Error rate of the null model:

```{r}
wells |> summarize(null_switch = mean(switch))
```
So the error rate is 1- that, about 42.5%

For the fitted model:

```{r}
pred <- predict(fit_a, type='response')
mean((pred > 0.5 & wells$switch==1) | (pred <= 0.5 &  wells$switch==0))
```

ok, so better then null!

### (e) Large Binns 

Create indicator variables corresponding to dist < 100; dist between 100 and 200; and
dist >  200. Fit a logistic regression for Pr(switch) using these indicators. With this new
model, repeat the computations and graphs for part (a) of this exercise.

```{r}
wells <- wells |> 
  mutate(
    dist_cut = cut(dist,c(0,100,200,1000))
 
  )

fit_e <- stan_glm(switch ~ dist_cut - 1 ,
                  data = wells, family= binomial(link="logit"), refresh = 0)
```

```{r}
print(fit_e, digits =3)
```

 

Plot it


```{r}
binned <- wells |> 
          group_by(dist_cut) |> 
          summarise(p_cut = mean(switch), sd_cut = sd(switch) / sqrt(n()) ) 
binned <- binned |>
  mutate(.pred = predict(fit_e, type = "response", newdata = binned))

binned |> 
  ggplot() +
  geom_point(aes(x=dist_cut, y= p_cut)) + 
  geom_errorbar(aes(x= dist_cut, ymin = p_cut - sd_cut,ymax = p_cut + sd_cut)) +
  geom_point(aes(x = dist_cut, y = .pred), color='red')
```

```{r}
binned
```

At first I was surprised, but then I recalled that fitting to a constant is the same as estimating the proportion, which is all we are doing for each bin! 

```{r}
as_tibble(fit_e) |> pivot_longer( everything(), names_to = 'bin', values_to = 'coef' ) |>
  group_by(bin) |> summarize(mean = mean(invlogit(coef)), sd = sd(invlogit(coef)))
```

### Alternate plot to better illustrate the fit IMHO

```{r}
binned <- wells |>  mutate(bin = cut_interval(dist, n = 10)) |>
        group_by(bin) |> 
             summarize(
             dist_mean = mean(dist),
             p_switch = mean(switch),
             p_se = sd(switch) / sqrt(n()) 
             )
             
 

binned |> 
  ggplot() +
  geom_errorbar(aes(x= dist_mean, ymin = p_switch - p_se,ymax = p_switch + p_se)) +
  geom_line(aes(x = dist, y = predict(fit_e, type = "response")),data=wells, color='red') +
  geom_point(aes(x=dist_mean, y= p_switch))
```

### How bad is the prediction accuracy?

```{r}
pred <- predict(fit_e, type='response')
mean((pred > 0.5 & wells$switch==1) | (pred <= 0.5 &  wells$switch==0))
```

Slightly better?
