# Chapter 5 selected exercises

## 5.2

```{r}
log_men <- 5.13
sd_men <- 0.17
log_women <- 4.96
sd_women <- .20

random_group <- function(size=10)
{
  men <- rbinom(size,1,0.5)
  weights <- ifelse(men, rlnorm(size,log_men,sd_men),rlnorm(size,log_women,sd_women))
  sum(weights)
}

weight_sim <- replicate(100000, random_group(10))
mean(weight_sim > 1750)
```

## 5.3

a) 
```{r}
dbinom(3,10,.4)
```

b)
```{r}
simulate_shots <- function(size = 10){
  rbinom(1,size,.4)
}

mean(replicate(10000, simulate_shots())==3)
```

## 5.4 

The mean of a single uniform deviate over range 0-1 is 0.5, with variance 1/12 (see previous chapter).
For 20 of them, the mean is 20*0.5= 10, and the variance would be 20/12= 5/3

```{r}

sim_n_uniform <- function(n=20){
  sum(runif(n,0,1))
}

simulated_sums <- replicate(1000,sim_n_uniform())

hist(simulated_sums, freq=FALSE)
x <- 0:150/10
lines(x,dnorm(x,mean=0.5*20, sd = sqrt(20)/sqrt(12) )  )

```
Seems like a good approximation to me


## 5.6

```{r}
savings_mean <- 5
savings_err <- 4
market_mean <- 40000
market_err <- 10000

simulated_market <- function(){
  savings <- rnorm(1,savings_mean,savings_err)
  market <- rnorm(1,market_mean,market_err)
  savings*market
}

data <- replicate(10000, simulated_market())
hist(data/1000)
mean(data)

```


## 5.7 Did this in chapter 4 presentation
 
## 5.8 

a)
```{r}
true_effect <- 0.10
sd_err <- 0.17

simulated_exp <- rnorm(1000, true_effect, sd_err)
```

b)

```{r}
lower= simulated_exp-2*sd_err
upper = simulated_exp + 2*sd_err

mean(lower < true_effect & upper > true_effect)
```

c)
```{r}
c(mean(simulated_exp),sd(simulated_exp))

```

Nothing here is suprising, not sure what the point is. 

```{r}
mean(simulated_exp > .35)
```
Note that 7 % of the point estimates are larger then 0.35


## 5.9

Ah, here is the point of 5.8!

```{r}
significant <- simulated_exp[lower>0]
c(mean(significant), sd(significant))
```

The mean is much higher , and the sd lower. This demostrates type M error.

## 5.10
a)

```{r}

cost_increase <- 600
cost_err <- 400

effect <- 3.0
effect_err <- 1.0

sim_cost <- rnorm(1000,cost_increase, cost_err)
sim_effect <- rnorm(1000,effect, effect_err)

plot(sim_cost,sim_effect)
```

b)
```{r}
cost_effect_ratio = sim_cost/sim_effect
quantile( cost_effect_ratio , probs = c(0.025,0.25,0.75,.975))
```
c)
```{r}
sim_effect <- rnorm(1000,effect, 2)
cost_effect_ratio = sim_cost/sim_effect
quantile( cost_effect_ratio , probs = c(0.025,0.25,0.75,.975))
```

## 5.12

```{r}
assign <- function(n,p=0.5, method = "independant"){
switch(method,
 independant = rbinom(n,1,p),
 complete = {
   ones <- rep(1,n*p)
   zeros <- rep(0,n-length(ones))
   sample(c(ones,zeros),n)
 },
 matched = {
   if(p!=0.5 || n%%2 != 0) stop("matched only works with p=0.5 and even n")
   pairs <- rbinom(n/2,1,0.5)
   res<- c()
   for(pair in pairs){
     if(pair)
       res <- c(res,c(0,1))
     else
       res <- c(res,c(1,0))
   }
   res
 },
 stop("Method must be one of independant, complete, or matched")
)
}

test <- assign(100,method="matched")  

```
 

