# Selected Exercises
```{r}
library(tidyverse)
library(rstanarm)
library(ggplot2)
library(rstan)
library(brms)
```

## 12.6

a) Create a scatterplot of mortality rate versus level of nitric oxides. Do you think linear regression will fit these data well? Fit the regression and evaluate a residual plot from the regression.

```{r}
pollution <- read_csv("Pollution/data/pollution.csv",
                      show_col_types =  FALSE)

ggplot(data = pollution, aes(x=nox, y = mort)) + geom_point()
```
I doesn't appear that a linear model will work very well.  Lets see:

```{r}
fit <- stan_glm(mort ~ nox, data = pollution, refresh =0)
fit
```

```{r}
ggplot(data = pollution, aes(x=nox, y = mort)) + geom_point() +
  geom_abline(intercept= fit$coefficients[1], slope = fit$coefficients[2])
```

Plotting the residuals:
```{r}
pollution$residual <- pollution$mort - predict(fit)
ggplot(data = pollution) + geom_point(aes(x=nox, y= residual))
```

Basically explains none of the variance.

b) I suspect that the log of the nox will look more linear, and to make it easier to interpret lets also do the log of the mort.

```{r}
ggplot(data = pollution, aes(x=log(nox), y = log(mort))) + geom_point()

```

```{r}
fit2 <- stan_glm(log(mort) ~ log(nox), data = pollution, refresh =0)
print(fit2, digits =4)
```



```{r}
ggplot(data = pollution, aes(x=log(nox), y = log(mort))) + geom_point() +
  geom_abline(intercept= fit2$coefficients[1], slope = fit2$coefficients[2])
```


Plotting the residuals:
```{r}
pollution$residual <- log(pollution$mort) - predict(fit2)
ggplot(data = pollution) + geom_point(aes(x=log(nox), y= residual))
```

Not convinced that this is much better.

```{r}
loo1 = loo(fit)
loo2 = loo(fit2)
loo2_with_jacobian <- loo2
# just copied from the book.... 
loo2_with_jacobian$pointwise[,1] <- loo2_with_jacobian$pointwise[,1] - log(pollution$mort)
loo_compare(loo1,loo2_with_jacobian)
```

Fit two is only barely better. 

That outlier is the pulling the fit, a later chapter will show how to do robust fit using brm, but i could not get this working on windows. 


```  
fit3 <- brm(mort ~ log(nox), data = pollution, family = 'student')
 
```

c) How do we interpret the coefficent on the slope?

The model is: `log(mort) = 0.016*log(nox) + intercept`

or `mort = exp(intercept)*nox^(0.016)`

The slope here give the exponent in the power law.

 
