## Inflation analyis

### 1.10

The goal of this project is to attempt to understand inflation in terms of the changes in the underlying monetary supply as well as effects of federal reserve interest rate targets. I also want to look at relationships between rate sof change of the various money measures, and how this is driven by federal reserve decisions and goverment defecit spending.

The primary data source for this will be the FRED service of the federal reserve.  Initial data to look at:

   - M1 -  note there was a change in May 2020, moving savings deposits into M1 (from M2) .  We intend to look at rates of change of this, we will have a spike to deal with.  Perhaps there is a real effect (this was due to a policy change removing limits on transfers from savings) - M1SL (Note we might be able to correct for spice by regressing rate on M1 vs M2)

  - M2  M2SL

  - Monetary Base  BOGMBASE (Note: NOT seasonally adjusted)
  
  - Inflation rate Issue here is change in weights. I intend to  BLS CPI For Urban consumers All Items .  Seems to be the common inflator in use. on fred this is CPIAUCSL
  
  - Federal reserve funds rate. FEDFUNDS
  
Data for future use only: 

  - Government spending (Future) Quarterly only. W068RCQ027SBEA is total expendatures, W006RC1Q027SBEA is total tax receipts. Need to dig in more to see if these are really what we want. Going to start with this series: 
  
  - GDP - FUTURE. only available quarterly
  
  Money supply and inflation in OTHER countries
  
This data seems to be all available on a monthly bases and is seasonally adjusted.  
  
### 2.10

The primary issue of validity is the changes in definitions. We will need to be careful to look for outliers in rates of change that will signal a change. Another issue of validity is the changes to how inflation are measured, we will need to pick a good measure.

Another challenge here is that this is time series data and we dont expect instant reactions. We might need to look at yearly data instead.

Looking at data, I see that Jan 1 1959 is a common starting point we can use at first.

### 3.10  Most likely model would be linear. See below

### 4.11

Nothing specificly seems to fit for this, but i may revisit in future chapters.

## Import data

Note that this data was read from Fred (using saved set "inflationResearch") and then converted from excel to csv. This process I need to automate. Data is monthly.

```{r}
library(tidyverse)
monetary_data <- read_csv('inflationResearch.csv')
```

I want to really look at differences, and at first i will restrict to 1980 and following so that I will hopefully have not so many changes in defitions to deal with.

```{r}
 
annual_growth <- \(monthlies)
{
  monthly_rate <- (monthlies - lag(monthlies))/lag(monthlies)
  100*(exp(log(1+monthly_rate)*12) - 1)
}

growth_data <- monetary_data |> 
        mutate(inflation = annual_growth(CPIAUCSL),
               base_growth = annual_growth(BOGMBASE),
               M2_growth = annual_growth(M2SL),
               M1_growth = annual_growth(M1SL)) |>
        select(-c(BOGMBASE, M1SL, M2SL,CPIAUCSL)) |> 
        filter(DATE >= '1980-01-01')
tail(growth_data)
```

We are aware of one big outlier in M1 Growth when the definition was changed.

```{r}
growth_data |> filter(DATE > '2020-01-01' & DATE < '2020-12-01')
```
## Initial exploration

```{r}
growth_data <- growth_data |> mutate(M1_growth = if_else(M1_growth > 100, NaN, M1_growth))

long_growth <- growth_data |>   pivot_longer(!DATE, names_to = "growth_type", values_to = "growth") #for plotting
 
```

```{r}
 
ggplot(long_growth, aes(x=DATE)) +
  geom_point(mapping = aes( y =  growth , col = growth_type))  
  
```

Note large outliers we still will need to deal with..


```{r} 
ggplot(long_growth |> filter(growth < 100), aes(DATE, growth)) + 
  geom_line() + 
  facet_wrap(~growth_type, scales = "free_y", ncol = 1)
```
 

## Monetary measures compared



For this first glance i will remove large outliers as above

```{r}
GGally::ggpairs(growth_data |> select(-c(DATE,inflation, FEDFUNDS)) |> filter(M1_growth < 50 & base_growth < 50))
```

```{r}
ggplot(data = growth_data |> filter(M1_growth < 50)) +
  geom_point(mapping = aes( x = M1_growth, y= M2_growth), col = 'blue') 
  
```
Looks like a linear model can work here for M2 vs M1 growth. 

 
However Base money seems to be poor predictor of M1  or M2 

## Inflation vs money supply

```{r}
inf_growth <- growth_data |> select(-DATE) |>   pivot_longer(!inflation, names_to = "growth_type", values_to = "growth") #for plotting

ggplot(inf_growth |> filter(growth < 50), aes(growth, inflation)) + 
  geom_point() + 
  facet_wrap(~growth_type)
```
 
 
None of the monetary measures seem to be very good at predicting inflation , at least monthly. We need to try different things, possibly look at moving averages of monetary growth to capture the lag. This might be a bit out of scope.
 
 

## Monetary base outliers

Although I dont plan to use Monetary base, since it seems to have no predictive value, we do see there are some big outliers. Monetary base has not been redefined, but policy changes directly impact this quantity! For example, the one in April 2020 seems quite a large increase in the reserve balances, but it was due to fed paying interest on reserves .

```{r}
growth_data |> filter(base_growth > 200)

```
 

## Focus on M2 and lagged M2

```{r}
m2_lag_data <- growth_data |> 
  select(-c('M1_growth','FEDFUNDS','base_growth','DATE')) |>
  mutate(M2_lag1 = lag(M2_growth), M2_lag2 = lag(M2_lag1), M2_lag3 = lag(M2_lag2),
         M2_lag4 = lag(M2_lag3),M2_lag5 = lag(M2_lag4), M2_lag6 = lag(M2_lag5)) 

m2_lag_data |> 
  pivot_longer(!inflation, names_to = 'months_lag', values_to = 'growth') |>
  filter(growth < 20) |>
  ggplot() +
  geom_point(mapping = aes(growth, inflation))+
  facet_wrap(~months_lag) 

```
### M2 growth by itself.


```{r}
m2_lag_data |>
ggplot() +
  geom_point(mapping = aes(M2_growth, inflation))
 
```


```{r}
library(rstanarm)
fit_m2 = stan_glm(inflation ~ M2_growth, data = m2_lag_data )
fit_m2$stanfit
```


```{r}
as.tibble(fit) |>  ggplot() + geom_density(mapping = aes(x= M2_growth))
```
Ok, so why is the correlation negative? That seems suprising. If anything I would expect it to be consistent with zero.   

Looks like the large outliers are the cause ? 

I dont know, lets just see if we remove all outliers, altough I would prefer to model them, this is jsut exploratory.

```{r}
m2_lag_os <- m2_lag_data |> filter(M2_growth < 20 & inflation > -9)
m2_lag_os |> ggplot() + geom_point(mapping = aes(x=M2_growth, y = inflation))
```
From thsi graph it is clear that there should be no depenance.

```{r}
fit_m2_os= stan_glm(inflation ~ M2_growth, data = m2_lag_os )
fit_m2_os$stanfit
```

Removing the outliers suggest no dependance on M2 growth at all. I suspect that the small effect although statistically significant was not significant at all.


Attempt robust (student-t) regression?

```{r}
library(brms)
fit_m2 = brm(inflation ~ M2_growth,family = student(), data = m2_lag_data )
```
```{r}
fit_m2
```


### Include more of the lags

```{r}
library(rstanarm)
fit = stan_glm(inflation ~ ., data = m2_lag_data )
fit$stanfit
```


Although we are seeing statistically significant predictors, none of them seem to be actually significant. 



## Focus on M1 and lagged M1

```{r}
m1_lag_data <- growth_data |> 
  select(-c('M2_growth','FEDFUNDS','base_growth','DATE')) |>
  mutate(M1_lag1 = lag(M1_growth), M1_lag2 = lag(M1_lag1), M1_lag3 = lag(M1_lag2),
         M1_lag4 = lag(M1_lag3),M1_lag5 = lag(M1_lag4), M1_lag6 = lag(M1_lag5)) 

m1_lag_data |> 
  pivot_longer(!inflation, names_to = 'months_lag', values_to = 'growth') |>
  filter(growth < 20) |>
  ggplot() +
  geom_point(mapping = aes(growth, inflation))+
  facet_wrap(~months_lag) 

```

 ## Focus on base and lagged base

```{r}
base_lag_data <- growth_data |> 
  select(-c('M2_growth','M1_growth','FEDFUNDS','DATE')) |>
  mutate(base_lag1 = lag(base_growth),base_lag2 = lag(base_lag1), base_lag3 = lag(base_lag2),
        base_lag4 = lag(base_lag3),base_lag5 = lag(base_lag4), base_lag6 = lag(base_lag5)) 

base_lag_data |> 
  pivot_longer(!inflation, names_to = 'months_lag', values_to = 'growth') |>
  filter(growth < 20) |>
  ggplot() +
  geom_point(mapping = aes(growth, inflation))+
  facet_wrap(~months_lag) 

```



